---
title: "On the origin of citations"
author: "Emily Davenport"
date: "September 15, 2015"
output:
  html_document:
    fig_caption: yes
    fig_height: 8
    fig_width: 8
    highlight: espresso
    number_sections: yes
    self_contained: no
    theme: cerulean
    toc: yes
---

# Load the data

## using read.delim

```{r load_data}
counts_raw <- read.delim("data/counts-raw.txt.gz")
counts_norm <- read.delim("data/counts-norm.txt.gz")
```

# Data exploration

What's the distribution of authors in all articles of our data set?

```{r author_histogram, echo=FALSE, fig.cap="Figure 1: Number of Authors per Article"}
hist(counts_raw$authorsCount, main="Authors per paper", xlab="# authors")
```

```{r facebook_counts, echo=FALSE, fig.cap="Figure 2: Facebook data"}
hist(counts_raw$facebookShareCount, main="Facebook popularity", xlab="# shares", xlim=c(0, 100))
```

The average number of Facebook shares per paper in the data set is `r mean(counts_raw$facebookShareCount)`


## dplyr


```{r}
library("dplyr")
```

```{r}
research <- filter(counts_raw, articleType == "Research Article")
```

```{r}
research_2006 <- filter(research, year == 2006) 
nrow(research_2006)
```

```{r}
research_2006_fb <- filter(research, year == 2006, facebookCommentCount > 0)
nrow(research_2006_fb)
```

```{r}
research_2006_fb_tweet_disease <- filter(research, year == 2006, facebookCommentCount > 0 | backtweetsCount > 0, grepl("Infectious Diseases", plosSubjectTags))
nrow(research_2006_fb_tweet_disease)
```

```{r}
colnames(research)
```

```{r}
article_info <- select(research, doi, pubDate, journal, title, articleType, authorsCount)
colnames(article_info)
```
#dplyr is special and you can select columns with contiguous names
```{r}
article_info <- select(research, doi:authorsCount)
colnames(article_info)
```
#makes a new db of any column with "Count" except authorsCount and adding the last two
```{r}
metrics <- select(research, contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
colnames(metrics)
```

#Challenge 2
```{r}
research_filtered <- filter(research, year < 2009, pdfDownloadsCount > 1000, mendeleyReadersCount > 15, wosCountThru2011 < 10)
low_cite <- select(research_filtered, journal, title, year)
dim(low_cite)
```

#Chaining commands with dplyr


pipe character %>% which will directly take the output of the last command as the input for the next


```{r}
facebook_2006 <- research %>% filter(year ==2006) %>% select(contains("facebook"))
head(facebook_2006)

```

arrange, works similar to the function order

```{r}
research %>% arrange(desc(authorsCount), desc(wosCountThru2011)) %>% select(authorsCount, wosCountThru2011) %>%
        slice(1:10)
```


```{r}
test <- research %>% arrange(desc(wosCountThru2011)) %>% select(title) %>% slice(1:3)
```


```{r}
research %>% arrange(desc(wosCountThru2011)) %>% select(authorsCount, title, journal, plosSubjectTags) %>% slice(1:3)
```

## Now creating new data with dplyr

```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished / 7,
                                yearsSincePublished = weeksSincePublished / 52)
research %>% select(contains("Since")) %>% slice(1:10)
```

using summarize: maintains it as a dataframe 

```{r}
research %>% summarize(plos_mean = mean(plosCommentCount),
                       plos_sd = sd(plosCommentCount),
                       num = n())
```

Function groupBy makes some nested for loops unnecessary

```{r}
research %>% group_by(journal) %>% summarize(tweets_mean = mean(backtweetsCount))
```


Challenge 3

```{r}
tweets_per_journal <- research %>% group_by(journal) %>% 
        summarize(num = n(), tweets_mean = mean(backtweetsCount), SEM = sd(backtweetsCount) / sqrt(num)) 
        
```

# ggplot2 play time

```{r}
library("ggplot2")
```

```{r plot1}
p <- ggplot(data = research, mapping = aes(x = pdfDownloadsCount,
                                           y = wosCountThru2011))
p + geom_point()
```


```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                          y = wosCountThru2011)) +
        geom_point(aes(color = journal, size = authorsCount))
p
```


```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                          y = wosCountThru2011)) +
        geom_point(aes(alpha = daysSincePublished))
p
```


```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                          y = wosCountThru2011)) +
        geom_point(aes(color = journal)) +
        geom_smooth()
p
```


If you color by journal before geoms it will do the smooth line applied to each journal group
```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                          y = wosCountThru2011, 
                          color = journal)) +
        geom_point() +
        geom_smooth()
p
```

only use aes when you want something to apply to a variable of your data
```{r}
p <- ggplot(research, aes(x = daysSincePublished,
                          y = wosCountThru2011)) +
        geom_point(aes(color = journal), alpha = .5) +
        geom_smooth(aes(color = "red"))
p
```

cannot have a loess curve because of zeros 
```{r}
p <- ggplot(research, aes(x = pdfDownloadsCount,
                          y = wosCountThru2011)) +
        geom_point(aes(color = journal)) +
        geom_smooth()
p + scale_x_log10() + scale_y_log10()
```


```{r}
p <- ggplot(research, aes(x = log10(pdfDownloadsCount + 1),
                          y = log10(wosCountThru2011 + 1))) +
        geom_point(aes(color = journal)) +
        geom_smooth() + 
        scale_x_continuous(breaks = c(1, 3), labels = c(10, 1000)) +
        scale_y_continuous(breaks = c(1, 3), labels = c(10, 1000))
p
```

Changing color scale
```{r}
p <- ggplot(research, aes(x = log10(pdfDownloadsCount + 1),
                          y = log10(wosCountThru2011 + 1))) +
        geom_point(aes(color = journal)) +
        geom_smooth() + 
        scale_x_continuous(breaks = c(1, 3), labels = c(10, 1000)) +
        scale_y_continuous(breaks = c(1, 3), labels = c(10, 1000))
p + scale_color_grey()
```


```{r}
library("RColorBrewer")
display.brewer.all(type = "qual")
```


```{r}
p <- ggplot(research, aes(x = log10(pdfDownloadsCount + 1),
                          y = log10(wosCountThru2011 + 1))) +
        geom_point(aes(color = journal)) +
        geom_smooth() + 
        scale_x_continuous(breaks = c(1, 3), labels = c(10, 1000)) +
        scale_y_continuous(breaks = c(1, 3), labels = c(10, 1000))
p + scale_color_brewer(palette = "Dark2")
```


If you want to change the figure legend you can:
```{r}
p + scale_color_brewer(palette = "Dark2", labels = 1:7, name = "title")
```


Challenge
```{r}
p <- ggplot(research, aes(x = sqrt(pdfDownloadsCount),
                          y = sqrt(wosCountThru2011))) +
        geom_point(aes(color = journal)) +
        geom_smooth() + 
        scale_color_brewer(palette = "Accent")
p
```


Facets!
```{r}
p <- ggplot(research, aes(x = log10(pdfDownloadsCount + 1),
                          y = log10(wosCountThru2011 + 1))) +
        geom_point(aes(color = journal)) +
        geom_smooth() + 
        scale_x_continuous(breaks = c(1, 3), labels = c(10, 1000)) +
        scale_y_continuous(breaks = c(1, 3), labels = c(10, 1000))
p + facet_wrap(~journal, ncol = 2)
```


Facet Grid
```{r}
research <- mutate(research, immuno = grepl("Immunology", plosSubjectTags))
p <- ggplot(research, aes(x = log10(pdfDownloadsCount + 1),
                          y = log10(wosCountThru2011 + 1))) +
        geom_point(aes(color = journal)) +
        geom_smooth() + 
        scale_x_continuous(breaks = c(1, 3), labels = c(10, 1000)) +
        scale_y_continuous(breaks = c(1, 3), labels = c(10, 1000))
p + facet_grid(journal~immuno)
```


Because it is so common to alter the original data frame there is a short cut
```{r}
p %+% research + facet_grid(journal~immuno)
```
I think this says to make p again but using research (which has been updated) and the facet addition


Challenge
```{r}
research <- mutate(research, evolution = grepl("Evolutionary Biology", plosSubjectTags))
p %+% research + facet_grid(evolution~immuno)
```

Other Geoms
```{r}
p_box <- ggplot(research, aes(x = journal,
                              y = log10(wosCountThru2011 + 1))) +
  geom_boxplot() +
  scale_y_continuous(breaks = c(1, 3), labels = c(10, 1000))
p_box
```


```{r}
tweets_per_journal <- research %>%
  group_by(journal) %>%
  summarize(num = n(),
            mean = mean(backtweetsCount),
            sem = sd(backtweetsCount) / sqrt(num))
tweets_per_journal
```

Including stat = identity is important so it does the mean and not the count
```{r}
tweets_bar <- ggplot(tweets_per_journal, aes(x = journal, y = mean)) +
  geom_bar(stat = "identity") +
        geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1) +
        geom_text(aes(label = num), hjust = 0, vjust = 0)
tweets_bar
```


```{r}
tweets_point <- ggplot(tweets_per_journal, aes(x = journal, y = mean)) +
  geom_point() +
        geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1) +
        geom_text(aes(label = num), hjust = 0, vjust = 0)
tweets_point
```


Challenge
```{r}
tweets_per_journal_year <- research %>%
  group_by(journal, year) %>%
  summarize(num = n(),
            mean = mean(backtweetsCount),
            sem = sd(backtweetsCount) / sqrt(num))
tweets_per_journal_year
tweets_bar_j_y <- ggplot(tweets_per_journal_year, aes(x = journal, y = mean)) +
  geom_bar(stat = "identity") +
        geom_errorbar(aes(ymin = mean - sem, ymax = mean + sem), width = .1) +
        geom_text(aes(label = num), hjust = 0, vjust = 0) +
        facet_wrap(~year)

```

looks way better with the main title
```{r}
tweets_bar_j_y + labs(title = "Mean tweets per journal per year", x = "Journal", y = "Mean number of tweets") + theme_bw()

```




Challenge
```{r}
sqrtCount2011 <- research %>% sqrt(wosCountThru2011)
geom_density(researc$wosCountThru2011 ~ sqrtCount2011$wosCountThru2011)

```


If you want to change the theme of all subsequent plots
theme_set(theme_bw())
